---
title: "Introduction to ggplot2 Themes"
author: "Boyd Skelton (Energent Group)"
date: "April 5, 2016"
output: 
  ioslides_presentation:
    logo: common/logo_no_text.png
---

```{r setup, include=FALSE}
source("load.R")
knitr::opts_chunk$set(echo = TRUE)

```
## {.flexbox .vcenter}

"There is no accounting for taste"

## Default Theme

```{r, echo = FALSE}

p <- ggplot(data, aes(x=permit_year, y=total, fill=pad_factor)) +
  geom_bar(position="fill", stat="identity") +
  labs(
    title= "Eagle Ford Well Pad Distribution",
    x = "Permit Year",
    y = "Pad Count",
    fill = "Wells per Pad"
  );p

```

## Custom Theme

```{r, echo = FALSE}

p + theme_eg() +
    scale_fill_brewer(palette="Set1")

```

## ggplot Structure

* Plot - entire plot
* Panel - chart area
* Legend
* Axis 
* Strip - facets

## Plot

```{r}

p + theme(plot.background = element_rect(fill = "red"))

```


## Panel

```{r}

p + theme(panel.background = element_rect(fill = "red"))

```

## Legend

```{r}

p + theme(legend.background = element_rect(fill = "red"))

```


## Theme Basics

* `element_text()` - style text objects (titles, labels, etc.)
```{r}
plot.title      = element_text(face="bold", color="red", size=15)
```
* `element_line()` - style line objects (axes, grid lines, ticks, etc.)
```{r}
axis.ticks      = element_line(colour = "black")
```
* `element_rect()` - style rect objects (backgrounds, borders)
```{r}
panel.border    = element_rect(fill = NA, colour = "white")
```

## Theme Basics (Continued)

* `unit()` - change magnitude (margins)
```{r}
panel.margin     = unit(0.0, "lines")
```

To set something to nothing use `element_blank()`. For example, no border:

```{r}
panel.border     = element_blank()
```

## Implementing Your Own Theme

1. Create a user-defined function
2. Start with a base theme (`theme_bw`, `theme_gray`)
3. Add additional themes to override (modify) the styles
4. Add it to your plot with `+`

```{r theme}

# custom function
custom_theme <- function(base_size = 10, base_family = "Helvetica")
{
  # start with base theme
  theme_bw(base_size = base_size, base_family = base_family) + 
  theme(
    plot.title = element_text(face="bold", colour="red", 
                              size=15, hjust=0.5, vjust=2)
  )
}

```

## Making Your Own Theme (Continued)

```{r}
# make title red
p + custom_theme() # or use theme_set()

```

## Other Themes, Add-Ins, & Comments

* **ggThemes** - CRAN Package of several popular themes (WSJ, Tableau, Excel, etc.)
* **ggThemeAssist** - RStudio Add-in (first commit March 2016)

**Comments**

1. Be careful of older examples - several items deprecated/removed
2. Things are quickly evolving making many "hacks" unnecessary

## Helpful Links

**Theme Vignettes**

http://docs.ggplot2.org/dev/vignettes/themes.html

**ggThemes**

https://cran.r-project.org/web/packages/ggthemes/vignettes/ggthemes.html

**ggThemeAssist**

https://github.com/calligross/ggthemeassist
